*Prochlorococcus marinus* responses to light and oxygen

### 2024-02-22 Code/Manuscript_O2.Rmd
edits to results & discussion

Near the equator, photoperiod remains relatively constant at the ocean surface, approximately 12 hours (h) of daylight and 12 h of darkness throughout the year. The effective length of the photoperiod does, however attentuate with depth as dawn and dusk light at depth drops below the level needed for biological processes. As *P. marinus* potentially expands its temperature-permissive niches poleward into temperate regions  [@flombaumPresentFutureGlobal2013; @bartonAnthropogenicClimateChange2016], it will encounter more
pronounced seasonal variations in photoperiod regimes both at surface and at depth, with potentially complex effects upon viable growth niches @liDiatomGrowthResponses2017 @prezelinDielPeriodicityPhytoplankton1992. For example, @vaulotGrowthProchlorococcusPhotosynthetic1995 showed that *Prochlorococcus* replication of DNA occurs in the afternoon, while cell division occurs at night.

@@

# Results {.unnumbered}

@@

ProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4, SS120 and MIT9313.",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison culture growth experimental conditions are indicated by horizontal grey lines for depths approximating imposed peak Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^); and vertical grey lines for O~2~ (µM).",
                       "Point colours identify *Prochlorococcus* clade assignments",
                       "Data from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

The growth rates of *P. marinus* MED4, Clade HLI, increased with higher imposed PAR and longer photoperiods
(Figure \@ref(fig:MED4PAR)), under 250 µM O~2~. No growth was
observed under any imposed conditions under a 4 h photoperiod. MED4 was
able to grow under the lowest PAR treatment of 30 µmol photons m^-2^
s^-1^ but only when the blue light photoperiod was longer than 4 h and the white LED and red light photoperiod was longer than 8 h. Growth rate increased
with increasing photoperiods for all spectral wavebands when exposed to
PAR of 90 and 180 µmol photons m^-2^ s^-1^. MED4 grew fastest for each
spectral waveband under 16 h photoperiods. The maximum growth rate
(µ~max~) was
`r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 180 uE blue
light and 16 h photoperiod.

@@

Similar to the growth trends under 250 µM O~2~, MED4
maintained at 25 µM O~2~ showed the fastest growth rates when the
photoperiod was 16 h for each spectral waveband, across PAR levels
(Figure \@ref(fig:MED4PAR)). The 25 µM O~2~ growth experiments showed
generally similar growth rates as the 250 µM O~2~ growth experiments,
with the interesting exception that growth rates increased by 40 % under
25 µM O~2~ at PAR of 30 µmol photons m^-2^ s^-1^ and 16 h blue photoperiod and increased by 55 % under 16 h red photoperiod, in contrast to
the 250 µM O~2~ growth experiments. White LED light treatments grew
slower, if at all, under 25 µM O~2~, only growing under higher PAR of 90 µmol photons m^-2^ s^-1^ and 16 h photoperiod and 180 µmol photons m^-2^ s^-1^ and 12 and 16 h photoperiods.
As under 250 µM O~2~, the fastest growth rate observed for each spectral
waveband at each imposed PAR were greatest under the 16 h photoperiod.
The µ~max~ was
`r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "25", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^  achieved under 180 µmol photons
m^-2^ s^-1^ blue light and 16 h photoperiod. The 4 h photoperiod
experiments under white LED light were not performed because no growth
was achieved when grown under an 8 h photoperiod.

@@

XXX Fig letters do not match as now we have red GAMs. Below is results for Blue GAMsXXXX

We employed GAM to summarize and visualize the complex responses of growth to the interactions of peak PAR, oxygen, photoperiod and blue vs. red spectral bands

The GAM model in Figure \@ref(fig:MED4GAM) summarizes MED4 growth in response to red (A,B) or blue (C,D) peak PAR and photoperiod across 2 imposed oxygen concentrations.  Under 250 µM O~2~ MED4 achieved fastest growth rates above peak blue light of ~180 µmol photons m^-2^ s^-1^, and the longest photoperiod of 16, indicated by the 0.64 d^-1^ contour line representing the
90^th^ percentile of maximum achieved growth rate (Figure \@ref(fig:MED4GAM)C).  Growth decreased with decreasing photoperiod and peak PAR.  Under red light growth was generally slower but the pattern of growth responses to photoperiod and PAR was similar (Figure \@ref(fig:MED4GAM)A).  We again note the exclusion of MED4 from growth under 4 h photoperiod under both red and blue light (Figure \@ref(fig:MED4GAM)A,C). Under 25 µM O~2~ MED4 showed similar growth responses, but was excluded from both 4 and 8 h photoperiods XXX GAM changed after removing 1 bad run XXX and does not reflect true pattern at higher photoperiod and light MIrielle to fix.

@@

XXX Fig letters do not match as now we have red GAMs. Below is results for Blue GAMsXXXX

The GAM model in Figure \@ref(fig:SS120GAM) summarizes growth
responses of SS120 red (A) or blue (B) peak PAR and photoperiod,
across the 2 imposed oxygen concentrations. Under 250 µM O~2~,
Figure \@ref(fig:SS120GAM)A showed highest growth
rates achieved below PAR of 50 µmol photons m^-2^ s^-1^ and
photoperiods between 8 and 12 h, indicated by the contour line labeled
0.29 d^-1^ (representing the 90^th^ percentile of achieved growth rate). Note the disjunct regions of the plots, resulting from variable growth successes of SS120 under 250 µM O~2~.  Growth rate patterns under red light were similar, although somewhat faster.
In contrast, under 25 µM O~2~ and a photoperiod of 8 h SS120 exploited all blue peak PAR levels, achieving faster growth rates at a higher PAR of \~100 µmol photons
m^-2^ s^-1^, indicated by the contour line labeled 0.5 d^-1^
(representing the 90^th^ percentile of achieved growth rate), out pacing
the 90^th^ percentile fastest growth rates under 250 µM O~2~. Under red light and 25 µM O~2~ light SS120 grew across most conditions of peak PAR and photoperiod, achieving fastest growth under long photoperiods and peak PAR below 50 \~100 µmol photons
m^-2^ s^-1^.  Thus, the designation of SS120 as a LL strain is dependent upon the [O~2]. SS120 did not, however,  grow under tested conditions at 2.5 µM O~2~.

@@

The growth rates of *P. marinus* MIT9313, Clade LLIV increased with longer
photoperiods under low 30 µmol photons m^-2^ s^-1^ peak PAR, under 250 µM O~2~ (Figure
\@ref(fig:MIT9313PAR)).  Under intermediate 90 µmol photons m^-2^ s^-1^ peak PAR growth rates decreased with increasing blue light photoperiod. Blue light did not induce growth at 180 µmol photons m^-2^ s^-1^ peak PAR, while cells shows only marginal growth under white LED and red light under the 8 h photoperiod . The µ~max~
was
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 30 µmol photons
m^-2^ s^-1^ blue light and 16 h photoperiod.

@@

Under 2.5 µM O~2~  MIT9313 showed growth, particularly under blue LED light, albeit
generally slower than under the parallel 25 µM O~2~ experiments (Figure
\@ref(fig:MIT9313PAR)). Growth estimates did, however, show scatter among replicates, suggesting the 2.5 µM O~2~  was near the tolerance limit for growth of  MIT9313. Growth rates increased with longer photoperiods under blue
light treatments under peak PAR of 90 µmol photons m^-2^ s^-1^ . Growth 
for MIT9313 under PAR of 180 µmol photons m^-2^ s^-1^ and blue light
treatment decreased with increasing photoperiods with full growth inhibition
under a 16 h photoperiod. The red light peak PAR of 180 µmol
photons m^-2^ s^-1^ showed similar growth rates for 8 and 12 h
photoperiods. The µ~max~ was
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "2.5", Photoperiod == 4, Par_ue == 180) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 4 h blue light
photoperiod and PAR of 180 µmol photons m^-2^ s^-1^.  XXXXTHIS DOES NOT LOOK correct; 12 h, 90 uE photons blue light looks higher.

The 2.5 µM O~2~ white LED treatments under 4, 8 and 16 h photoperiods
and red light under 4 and 16 h photoperiods were not performed due to
time constraints.

@@

The GAM model in Figure \@ref(fig:MIT9313GAM) summarizes MIT9313 growth responses to red (A) or blue (B) peak PAR and photoperiod. Under 250 µM O~2~,  Figure \@ref(fig:MIT9313GAM)A shows MIT9313 achieves
fastest growth rates between blue peak PAR of 30 µmol photons m^-2^
s^-1^ and 50 µmol photons m^-2^ s^-1^ and photoperiods longer than 8 h,
indicated by the contour line labeled 0.51 d^-1^ representing the
90^th^ percentile of achieved growth rates. Figure
\@ref(fig:MIT9313GAM)A also shows that growth rate increases
with longer photoperiods, as long as the blue peak PAR levels remain below 50 µmol
photons m^-2^ s^-1^.  In contrast, under red light and 250 µM O~2~ MIT9313 grows faster while exploiting higher peak PAR and longer photoperiods.  Thus the designation of MIT9313 as a LL clade is dependent upon [O2].
(Figure \@ref(fig:MIT9313GAM)B).  Figure \@ref(fig:MIT9313GAM)C
shows that MIT9313 maintains growth even under 2.5 µM O~2~, under photoperiods between 4 and 8 h and peak blue PAR between 50 to 10 µmol photons
m^-2^ s^-1^ PAR (Figure \@ref(fig:MIT9313GAM)C).

@@

Cumulative diel absorbed PUR can potentially collapse photoperiod, PAR and spectral wavebands to a common metric of absorbed light. We plot observed growth rate vs. cumulative diel absorbed PUR to determine whether growth is a simple response to absorbed light per day across imposed photoperiods and spectral wavebands, or whether photoperiod or spectral wavebands have specific or interactive influences on growth of each ecotype. Cumulative daily absorbed PUR dose (µmol photons m^-2^ d^-1^) was calculated from the imposed absorbed PUR (µmol photons m^-2^ s^-1^) and photoperiod (h).

Due to the highly effective absorption of pigments in the blue spectral waveband range for all ecotypes, the maximum imposed cumulative diel absorbed PUR under blue light treatment is almost 3 times that of white LED light and about 6 times the red light treatment, despite being derived from the same imposed photoperiods and PUR regimes. As such, only growths rates under blue light extend beyond absorbed PUR of ~ 2 x 10^6^ µmol photons m^-2^ d^-1^.

@@

The representative of HLI clade, *P. marinus* MED4, showed no growth under any 4 h photoperiod treatments, even when a 4 h photoperiod delivered cumulative diel PUR
equivalent to other photoperiod treatments. In parallel MED4 showed no growth under 2.5 µM O~2~, no matter the level of diel cumulative absorbed PUR. 
In contrast, under 250 or 25  µM O~2~, and under any photoperiod greater than 4 h, MED4 growth was well described by a simple saturating response of growth to increasing cumulative diel absorbed PUR, with saturation of growth rate achieved around  1.0 x 10^6^ µmol m^-2^d^-1^ (Figure \@ref(fig:AllStrainPurFitsPlots)A).   XXXXDISTINCT PHOTOPERIOD FITSXXXXX

@@

The representative of the LLIII clade, *P. marinus* SS120 showed almost no growth under 2.5 µM O~2~ experiments. A majority of the 4 h photoperiod treatments did not grow under 250 µM O~2~ experiments even when a 4 h photoperiod delivered the cumulative diel PUR equivalent to other as other photoperiod treatments. SS120 did not grow when exposed to more than ~1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative diel PUR under any spectral waveband or photoperiod combination, under 250 µM O~2~ experiments(Figure \@ref(fig:AllStrainPurFitsPlots)B).

Under both 25 or 250 µM O~2~ experiments, SS120 growth plateaued by about 5.0 x 10^5^ µmol photons m^-2^ d^-1^ diel PUR, with some scatter among photoperiod and spectral waveband regimes. The onset of growth inhibition extended to higher cumulative diel absorbed PUR for cultures growing under 25 µM O~2~. XXXXDISTINCT PHOTOPERIOD FITSXXXXX

@@

The LLIV clade, *P. marinus* MIT9313, showed growth rising to a plateau by about 5 x 10^5^ µmol photons m^-2^ d^-1^ of cumulative diel PUR, with higher growth rates over a narrower plateau under 25 and 250 µM O~2~, compared to a wider, flatter response to cumulative diel  PUR under 2.5 µM O~2~. Above about 1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative PUR under 250 µM O~2~, MIT9313 showed full inhibition of growth, across photoperiods, and spectral wavebands.  In contrast, under 25 µM O~2~, MIT9313 showed a greatly extended exploitation of higher cumulative diel PUR, with full growth inhibition only above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^.  Similarly, under 2.5 µM O~2~, MIT9313 grew more slowly, but only showed full growth inhibition above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^ cumulative diel PUR (Figure \@ref(fig:AllStrainPurFitsPlots)C).  XXXXDISTINCT PHOTOPERIOD FITSXXXXX

@@

## Photosystem II maintenance as a limitation on *Prochlorococcus* growth {.unnumbered}

@@

Under full atmospheric [O~2~], LL clades of *Prochlorococcus* are restricted to growth under low light, in part because they suffer photoinhibition of Photosystem II (PSII) through several paths, including direct absorbance of UV or blue light, in parallel with generation of Reactive Oxygen Species (ROS) if the electron flow is slowed, [@aroPhotoinhibitionPhotosystemII1993], producing  damaging singlet oxygen (^1^O~2~) [@aroPhotoinhibitionPhotosystemII1993; @soitamoPhotoinhibitionMarinePicocyanobacteria2017; @murphyPhotoinactivationPhotosystemII2017; @hakalaEvidenceRoleOxygenevolving2005].
Repair of inhibited PSII process relies on the removal of damaged PsbA (Mann et al., 2000; Adam et al., 2005), followed by reassembly with newly synthesized PsbA [@nixonRecentAdvancesUnderstanding2010]. 
Thus, degradation of PsbA is a rate-limiting step in recovery from photoinhibition (Kanervo et al., 1993), mediated largely by a heterohexamer of (FtsH12)3, a membrane-bound (Sacharz et al., 2015)(Zak et al., 2001) metalloprotease  (Chiba et al., 2002) (Yoshioka-Nishimura and Yamamoto, 2014).  (Adam et al., 2005; Boehm et al., 2012) (Nixon et al. (2010)) (Komenda et al., 2007). (Pisareva et al., 2007; Sacharz et al., 2015),  (Mann et al., 2000), (AMANDA REF from ALGATECH).

>>>>

Near the equator, photoperiod remains relatively constant at the ocean surface, approximately 12 hours (h) of daylight and 12 h of darkness throughout the year. The effective length of the photoperiod does, however attentuate with depth as dawn and dusk light at depth drops below the level needed for biological processes. As *P. marinus* potentially expands its temperature-permissive niches poleward into temperate regions  [@flombaumPresentFutureGlobal2013; @bartonAnthropogenicClimateChange2016], it will encounter more
pronounced seasonal variations in photoperiod regimes both at surface and at depth, with potentially complex effects upon viable growth niches @liDiatomGrowthResponses2017 @prezelinDielPeriodicityPhytoplankton1992. For example, @vaulotGrowthProchlorococcusPhotosynthetic1995 showed that *Prochlorococcus* replication of DNA occurs in the afternoon, while cell division occurs at night.  To our knowledge, no study has as yet addressed *P. marinus* growth responses in relation to a  range of photoperiods.

@@

# Results and discussion {.unnumbered}

@@

ProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4, SS120 and MIT9313.",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison culture growth experimental conditions are indicated by horizontal grey lines for depths approximating imposed peak Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Point colours identify *Prochlorococcus* clade assignments",
                       "Data from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

The growth rates of *P. marinus* MED4, Clade HLI, increased with higher imposed PAR and longer photoperiods
(Figure \@ref(fig:MED4PAR)), under 250 µM O~2~, across all spectral wavebands. No growth was
observed under any imposed conditions under a 4 h photoperiod.The maximum growth rate
(µ~max~) was
`r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 180 uE blue light and 16 h photoperiod.

@@

Similar to the growth trends under 250 µM O~2~, MED4
maintained at 25 µM O~2~ showed the fastest growth rates when the
photoperiod was 16 h for each spectral waveband, across PAR levels
(Figure \@ref(fig:MED4PAR)).
The µ~max~ was `r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "25", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^  achieved under 180 µmol photons
m^-2^ s^-1^ blue light and 16 h photoperiod. The 4 h photoperiod
experiments under white LED light were not performed because no growth
was achieved when grown under an 8 h photoperiod of white LED light.

@@

The GAM model in Figure \@ref(fig:MED4GAM) summarizes MED4 growth responses to red (A,B) or blue (C,D) peak PAR and photoperiod across 2 imposed oxygen concentrations.  Under 250 µM O~2~ MED4 achieved fastest growth rates above peak blue light of ~180 µmol photons m^-2^ s^-1^, and the longest photoperiod of 16, indicated by the 0.64 d^-1^ contour line representing the
90^th^ percentile of maximum achieved growth rate (Figure \@ref(fig:MED4GAM)C).  Growth decreased with decreasing photoperiod and decreasing peak PAR.  Under red light growth was generally slower but the pattern of growth responses to photoperiod and PAR was similar (Figure \@ref(fig:MED4GAM)A).  Note the exclusion of MED4 from growth under 4 h photoperiod under both red and blue light (Figure . Under 25 µM O~2~ MED4 showed similar growth responses, but was excluded from both 4 and 8 h photoperiods.  MED4 did not grow under 25 µM O~2~, so no GAM model was run.

The representative of HLI clade, *P. marinus* MED4, was completely
excluded from growth under the 4 h photoperiod \@ref(fig:MED4GAM)A,C), and showed
increasing growth rates with increasing photoperiod. Thus, MED4 can
potentially exploit longer summer photoperiods in temperate zones, but
will likely remain excluded from growth in winter through its requirement for
more than 4 h of light per day, even if water temperature warms into the Clade HLI tolerance range.

MED4 grew under all PAR levels, so long as the photoperiod was longer than 4 h,
although growth rates were slow under the lowest PAR level of 30 µmol
photons m^-2^ s^-1^. Considering the range of PAR levels that MED4 can
utilize, it is reasonable to suggest that MED4 can inhabit not just
shallow depths, where light levels are high, but also deeper regions
characterized by lower light levels, subject to the limitation of a
photoperiod of at least 4 h even after depth attenuation of light. The photo
regimes of winter temperate zones, due to shorter photoperiods, exclude
MED4 from growth at all depths, however temperate photoperiods and light
levels for the remainder of the year are adequate to support growth.

MED4, grew under all tested spectral wavebands. Growth rates were
generally comparable under blue, white LED and red lights when exposed
to PAR levels greater than 30 µmol photons m^-2^ s^-1^ for each
photoperiod. At the lower PAR of 30 µmol photons m^-2^ s^-1^, MED4 grew
better under blue light. All growth rates for each spectral waveband
increased with increasing photoperiod. The ability to utilize all
spectral wavebands at high PAR levels and longer photoperiods make the
near surface temperate zones available niches for MED4. MED4 ability to
exploit blue light under low PAR levels and variable photoperiods
greater than 4 h allow MED4 to exploit deeper regions in these temperate
zones, where blue light prevails.

@@

The GAM model in Figure \@ref(fig:SS120GAM) summarizes growth
responses of SS120 to red (A) or blue (B) peak PAR and photoperiod,
across the 2 imposed oxygen concentrations. Under 250 µM O~2~,
Figure \@ref(fig:SS120GAM)A showed highest growth
rates below PAR of 50 µmol photons m^-2^ s^-1^ and
photoperiods between 8 and 12 h, indicated by the contour line labeled
0.29 d^-1^ (representing the 90^th^ percentile of achieved growth rate). The disjunct regions of the GAM plot results from variable growth success of SS120 under 250 µM O~2~.  Growth rate patterns under red light were similar, although somewhat faster.
In contrast, under 25 µM O~2~ and a photoperiod of 8 h SS120 exploited all blue peak PAR levels, achieving faster growth rates at a higher PAR of \~100 µmol photons
m^-2^ s^-1^, indicated by the contour line labeled 0.5 d^-1^
(representing the 90^th^ percentile of achieved growth rate), out pacing
the 90^th^ percentile fastest growth rates under 250 µM O~2~. Under red light and 25 µM O~2~ light SS120 grew across most conditions of peak PAR and photoperiod, achieving fastest growth under long photoperiods and peak PAR below 50 \~100 µmol photons
m^-2^ s^-1^.  Thus, the designation of SS120 as a LL strain is dependent upon the [O~2]. SS120 did not, however, grow reliably under tested conditions at 2.5 µM O~2~.

SS120, grew under a PAR of 30 µmol photons m^-2^ s^-1^, and under some conditions at PAR of 90 µmol photons m^-2^ s^-1^, but growth was generally inhibited under PAR of 180 µmol photons
m^-2^ s^-1^. These findings align with the results reported by
@mooreComparativePhysiologySynechococcus1995 and are expected for a low
light clade. Given that SS120 can tolerate longer photoperiods and low
PAR levels, the deep ocean temperate zones are a viable habitat but not
during temperate winter due to shorter photoperiods.

SS120 exhibited different growth responses to spectral wavebands
under varying O~2~ conditions, as such, this section will address the
growth response under 250 µM O~2~ experiments, while the interactions of
spectral wavebands under all O~2~ treatments will be discussed in the
next section. SS120, grew under all spectral wavebands at a PAR of 30
µmol photons m^-2^ s^-1^. Growth was completely inhibited under blue
light at PAR levels of 90 and 180 µmol photons m^-2^ s^-1^, but growth
occurred at PAR of 90 µmol photons m^-2^ s^-1^ under all other spectral
wavebands. Generally, growth was strongly inhibited under PAR of 180
µmol photons m^-2^ s^-1^ for all spectral wavebands. The ability of SS120 to grow under low blue light and longer photoperiods, makes the deep ocean waters of the temperate zone a viable potential habitat. Under 250 µM O~2~ SS120 is constrained to deeper ocean waters through its intolerance of higher blue PAR levels.

@@

The growth rates of *P. marinus* MIT9313, Clade LLIV increased with longer
photoperiods under low 30 µmol photons m^-2^ s^-1^ peak PAR, under 250 µM O~2~ (Figure
\@ref(fig:MIT9313PAR)).  Under intermediate 90 µmol photons m^-2^ s^-1^ peak PAR growth rates decreased with increasing blue light photoperiod. Blue light did not induce growth at 180 µmol photons m^-2^ s^-1^ peak PAR, while MIT9313 showed only marginal growth under white LED and red light under the 8 h photoperiod . The µ~max~ was `r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 30 µmol photons m^-2^ s^-1^ blue light and 16 h photoperiod.

@@

MIT9313 grew under 2.5 µM O~2~  particularly under blue LED light, albeit
generally slower than under the parallel experiments at 25 µM O~2~  (Figure
\@ref(fig:MIT9313PAR)). Growth estimates did show scatter among replicates, suggesting  2.5 µM O~2~  is near the tolerance limit for growth of MIT9313. Growth rates increased with longer photoperiods under blue
light treatments under peak PAR of 90 µmol photons m^-2^ s^-1^ . Growth  for MIT9313 under PAR of 180 µmol photons m^-2^ s^-1^ and blue light treatment decreased with increasing photoperiods with full growth inhibition under a 16 h photoperiod. The red light peak PAR of 180 µmol photons m^-2^ s^-1^ showed similar growth rates for 8 and 12 h photoperiods. The µ~max~ was `r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "2.5", Photoperiod == 4, Par_ue == 180) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 4 h blue light photoperiod and PAR of 180 µmol photons m^-2^ s^-1^.  XXXXTHIS DOES NOT LOOK correct; 12 h, 90 uE photons blue light looks higher. The 2.5 µM O~2~ white LED treatments under 4, 8 and 16 h photoperiods and red light under 4 and 16 h photoperiods were not performed due to time constraints.

@@

The GAM model in Figure \@ref(fig:MIT9313GAM) summarizes MIT9313 growth responses to red (A) or blue (B) peak PAR and photoperiod. Under 250 µM O~2~, Figure \@ref(fig:MIT9313GAM)A shows MIT9313 achieves
fastest growth rates between blue peak PAR of 30 µmol photons m^-2^
s^-1^ and 50 µmol photons m^-2^ s^-1^ and photoperiods longer than 8 h,
indicated by the contour line labeled 0.51 d^-1^ representing the
90^th^ percentile of achieved growth rates. Figure
\@ref(fig:MIT9313GAM)A also shows that growth rate increases
with longer photoperiods, as long as the blue peak PAR levels remain below 50 µmol
photons m^-2^ s^-1^.  In contrast, under red light and 250 µM O~2~ MIT9313 grows faster while exploiting higher peak PAR and longer photoperiods.  Figure \@ref(fig:MIT9313GAM)C
shows that MIT9313 maintains growth even under 2.5 µM O~2~, under photoperiods between 4 and 8 h and peak blue PAR between 50 to 10 µmol photons m^-2^ s^-1^ PAR (Figure \@ref(fig:MIT9313GAM)C).  Thus the designation of MIT9313 as a LL clade is dependent upon [O2] and light spectra. (Figure \@ref(fig:MIT9313GAM)B).

MIT9313, sometimes grew under all PAR levels but generally exhibited slower growth rate under PAR of 180 µmol photons m^-2^ s^-1^ and then only under some conditions. MIT9313 is well suited to inhabit the deep ocean in temperate zones with wide tolerance for photoperiod yet can also tolerate depths closer to the surface through its partial tolerance of higher PAR levels.

MIT9313, under 250 µM O~2~ grew under all spectral wavebands when exposed to PAR of 30 µmol photons m^-2^ s^-1^ and 90 µmol photons m^-2^ s^-1^ and was even able to grow under PAR of 180 µmol photons
m^-2^ s^-1^ for all spectral wavebands except blue light. MIT9313 can
thus exploit blue light in deeper waters of temperate zones and can even
tolerate moderate depths to exploit higher blue light levels, as long as
the photoperiod is shorter than 12 h.

@@

Cumulative diel absorbed PUR can potentially collapse photoperiod, PAR and spectral wavebands to a common metric of photosynthetically active light absorbed per day. Cumulative daily absorbed PUR dose (µmol photons m^-2^ d^-1^) was calculated from the imposed absorbed PUR (µmol photons m^-2^ s^-1^) and photoperiod (h). We plotted growth rates vs. cumulative diel absorbed PUR to determine whether growth is a simple response to diel PUR, across imposed photoperiods and spectral wavebands, or whether photoperiod or spectral wavebands have specific or interactive influences on growth.

Due to the absorption of *P. marinus* pigments in the blue spectral waveband range, the maximum imposed cumulative diel absorbed PUR under blue light is almost 3 times that of white LED light, and about 6 times the red light treatment, despite being derived from the same imposed photoperiods and peak PAR regimes. As such, only blue light experiments extend beyond cumulative diel PUR of ~ 2 x 10^6^ µmol photons m^-2^ d^-1^.

@@

The representative of HLI clade, *P. marinus* MED4, showed no growth under any 4 h photoperiod treatments, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments. In parallel MED4 showed no growth under 2.5 µM O~2~, no matter the level of diel cumulative PUR. 
In contrast, under 250 or 25  µM O~2~, and under any photoperiod greater than 4 h, MED4 growth was well described by a saturating response (REFERENCE for EQUATION) of growth to increasing cumulative diel PUR, with saturation of growth rate achieved around  1.0 x 10^6^ µmol m^-2^d^-1^ (Figure \@ref(fig:AllStrainPurFitsPlots)A), and no evidence of inhibition of growth at any achieved cumulative diel PUR.   XXXXDISTINCT PHOTOPERIOD FITSXXXXX

@@

The representative of the LLIII clade, *P. marinus* SS120 showed almost no growth under 2.5 µM O~2~ experiments. A majority of the 4 h photoperiod treatments did not grow under 250 µM O~2~ experiments, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments. SS120 did not grow when exposed to more than ~1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative diel PUR under any spectral waveband or photoperiod combination, under 250 µM O~2~ experiments(Figure \@ref(fig:AllStrainPurFitsPlots)B).

Under both 25 or 250 µM O~2~ experiments, SS120 growth plateaued by about 5.0 x 10^5^ µmol photons m^-2^ d^-1^ diel PUR, with some scatter among photoperiod and spectral waveband regimes. The onset of growth inhibition extended to higher cumulative diel PUR for cultures under 25 µM O~2~, showing that SS120 is partially protected from photoinhibition of growth by 25 µM O~2~. XXXXDISTINCT PHOTOPERIOD FITSXXXXX

@@

The LLIV clade representative, *P. marinus* MIT9313, showed growth rising to a plateau by about 5 x 10^5^ µmol photons m^-2^ d^-1^ of cumulative diel PUR, with higher growth rates over a narrower plateau under 25 and 250 µM O~2~, compared to a wider, flatter response to cumulative diel PUR under 2.5 µM O~2~. Above about 1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative PUR under 250 µM O~2~, MIT9313 showed full inhibition of growth, across photoperiods, and spectral wavebands.  In contrast, under 25 µM O~2~, MIT9313 showed a greatly extended exploitation of higher cumulative diel PUR, with full growth inhibition only above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^.  Similarly, under 2.5 µM O~2~, MIT9313 grew more slowly, but only showed full growth inhibition above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^ cumulative diel PUR (Figure \@ref(fig:AllStrainPurFitsPlots)C).  XXXXDISTINCT PHOTOPERIOD FITSXXXXX

@@

## Photosystem II maintenance, oxygen metabolism, and DNA repair as limitations on *Prochlorococcus* growth {.unnumbered}

@@

Under full atmospheric [O~2~] and bluelight, LL clades of *Prochlorococcus* are restricted to growth under low light, in part because they suffer photoinhibition of Photosystem II (PSII) through several paths, including direct absorbance of UV or blue light, in parallel with generation of Reactive Oxygen Species (ROS) if the electron flow is slowed, [@aroPhotoinhibitionPhotosystemII1993], producing  damaging singlet oxygen (^1^O~2~) [@aroPhotoinhibitionPhotosystemII1993; @soitamoPhotoinhibitionMarinePicocyanobacteria2017; @murphyPhotoinactivationPhotosystemII2017; @hakalaEvidenceRoleOxygenevolving2005].
Repair of inhibited PSII process relies on the removal of damaged PsbA (Mann et al., 2000; Adam et al., 2005), followed by reassembly with newly synthesized PsbA [@nixonRecentAdvancesUnderstanding2010]. 
Degradation of PsbA is a rate-limiting step in recovery from photoinhibition (Kanervo et al., 1993), mediated largely by a heterohexamer of (FtsH12)3, a membrane-bound (Sacharz et al., 2015)(Zak et al., 2001) metalloprotease  (Chiba et al., 2002) (Yoshioka-Nishimura and Yamamoto, 2014).  (Adam et al., 2005; Boehm et al., 2012) (Nixon et al. (2010)) (Komenda et al., 2007). (Pisareva et al., 2007; Sacharz et al., 2015),  (Mann et al., 2000), (AMANDA REF from ALGATECH).

*Prochlorococcus* genomes encode 4 FtsH proteins, henceforth referred to as FtsH1-4, homologs to the characterized FtsH isoforms of the model cyanobacterium *Synechocystis* FtsH, and with presumably parallel functions (table \@ref(tab:FtsHTable)). Upon a shift to higher light HLIII MED4 upregulated expression of FtsH1 and FtsH2 [@bonisteelStrainSpecificDifferences2018], homologs to the *Synechocystis* slr0228 and slr1604, implicated in PSII repair. In contrast representative LLIV strain MIT9313 showed less overall expression of the FtsH proteases, and thus has fewer FtsH serving each photosystem.  Furthermore MIT9313 expressed primarily FtsH3, homologous to Sll1463 involved in PSI biogenesis, and FtsH expression did not increase in response to light stress. Through adaptation to steady low light, LLIV *Prochlorococcus* instead allocates resources processes other than dynamic regulation of PSII repair.
